Detecting Spam on Social Networking Sites

### 2012-05-08 docs/paper/paper.tex
update

Heymann et al. \cite{heymann} survey the field of spam on SNSs, identifying several common approaches. Identification-based approaches identify spam to train classifiers based on labels submitted by users or trusted moderators. Rank-based approaches demote visibility of questionable content, while interface-based approaches apply policies to prevent unwanted behavior. This work groups classification-based approaches with detection, although classifiers can be used in conjuction with user information to prevent spam before it happens.

@@

De Wang et al. \cite{wang} propose a cross-site spam detection framework to share spam data across all social networking sites, building classifiers to identify spam in profiles, messages and web pages. This multi-pronged approach lends itself to associative classification, in which, for example, a message would be classified as spam if it contained a link to a web page that had a high probability of being spam. Unfortunately, the differing characterestics of various social networks \eg the length of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora across diverse sites.

Facebook \cite{stein} provides an overview of their ``immune system" defences against phishing, fraud and spam. The system is composed of classifier services, an ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare features for classification and a policy engine to take action on suspected misuse. While the discussion remains high-level and includes few implemention particulars, it does include significant detail on the various types and characterists of undesirable activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.

@@

Bosma et al. \cite{bosma} explore user-generated spam reports as a tool for building an unsupervised spam detection framework for SNSs. Their approach counts the number of spam reports against a suspected spammer and adds weight to reports based on user reputation. Determining reputation and trustworthiness of users in social networks has been well studied \cite{bian, guha, zhang} and appears to be a promising addition to social spam classification. The framework uses a Bayesian classifier and links messages with similar content, but does not take into account other features. Nevertheless, this is one of the few studies to test its framework on non-public data, inlcuding private messages, spam reports and user profiles from a large Dutch social networking site.

@@

For this project, we had unlimited access to data from InterPals, a SNS for users who wish to communicate with others from other countries, whether for language practice, cultural exchange or friendship. Users sign up by completing a registration form with information about themselves, including age, sex and location. After registering, users can expand their profile page to include descriptions, interests, languages they speak, etc. and upload photographs. After clicking on a link sent to their email addres, a user can begin to interact with others on the site via private message, public ``wall" posts, comments on photos and on a bulletin board system.

@@

\textbf{Message Body \& Subject:} We extracted both the message body and subject from 
each spam and ham message. The cleaning script generated two additional features based 
on this. First, we counted the frequency of non-standard punctuation (we noticed that many 
spam messages would, for example, put spaces before commas, periods and quotation marks, 
and omit spaces after these characters). Second, we calculated the ratio of uppercase to 
lowercase letters in the text, after observing relatively high amounts of uppercase text in spam messages.
%% add numbers

\textbf{Recipient Age/Sex/Country:} These fields were the age, sex and ISO 3166-1 alpha-2 code of the message 
recipient, as listed on their profile. Unfortunately, due to data collection issues, this data was only available 
for the most recent 11,000 spam messages. While the male-to-female ratio of the recipients was very close to that 
of non-spam recipients (both were 47\% male to 53\% female), we found that the age of recipients was typically higher.

\textbf{Recipient Replied:} A boolean, 1 or 0, depending on whether the recipient replied to the message. We saw 
that the mean reply rate for ham messages was 81.77\% (SD: 38.61), while it was only 0.81\% (SD: 8.97) for spam messages.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52, median 2. Ham: mean 175.51, std. dev. 322.19, median 64}
    \label{fig:uniqip}
\end{figure}

@@

% Modified from Cliff & Antonio's HW2 write-up
%Linear regression is a method to model the relationship between an output variable, $y$, and explanatory variables, $X$. The output variable is a linear sum of the explanatory variables multiplied by their corresponding coeffecients, represented by $\beta$. In our case, $y$ represents a review's ranking from one to five, while $X$ is the bag-of-words representation of the review text. This implies that $\beta$ represents the weight of a word's correlation to a higher review. This gives us the simple regression formula: $$y=X\beta$$
%
%In this project, we performed ordinary least squares regression. By minimizing the sum of squared residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution. $$\beta=(X'X)^{-1}X'y$$ In order to ensure that the matrix $X'X$ is invertible, we perform ridge regression. In ridge regression, we add the identity matrix, $I$, scaled by a factor, $\lambda$, to produce the following equation. $$\beta=(X'X+\lambda I)^{-1}X'y$$ We calculate this $\beta$ for a training set of review text along with their corresponding ratings which range from 1 to 5. We then evaluate our model $\beta$ on the withheld reviews in order to perform a 10-fold cross-validation.
%
%All of our linear regression code was implemented in MATLAB because of its ease-of-use and optimizations on matrix multiplications. We parsed the reviews from the tokenized binary format into a bag-of-words representation stored as a sparse matrix. Matlab uses Compressed Sparse Column format, so we actually stored the matrix $X'$ because the dictionary is very sparse. We also filtered out duplicate reviews by hashing the first ten words in a review to determine distinctness. This filtered the one million reviews to 492,130. 
%
%The actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter. The multiplication $X'X$ takes up too much memory if we use the entire dictionary, so we chose to limit the dictionary size. We found that a dictionary of the 10,000 most frequent words gave sufficient features while also being computable using MATLAB's built-in sparse matrix multiplications. We also tried using a dictionary of 15,000 words, but only had a marginal improvement along with much slower execution. Using our 10,000 word dictionary, we calculated the performance of our matrix multiply to be 0.123 Gflops/s by dividing the number of non-zero elements in our matrices by the time to compute the multiplication. 
%
%
%Since we used ridge regression, we had the ability to tune the $\lambda$ parameter in order to get the best performance. As shown in figure \ref{fig:lambda}, we observed an AUC improvement of about .5\% between the default value of 1 and a tuned parameter of 1500. This is a fairly major difference since the AUC is at 93\% using the default value. Additionally, we also noticed a major difference in the top words when using better $\lambda$ values. Using the default value of 1, our top words included obscure words including ``\_c'', ``\_j'', ``donovan'', and ``matlock.'' When we used the tuned $\lambda$ value of 1500, the words were much more representative, as shown in table \ref{tab:words}.
%

@@

Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is a real-valued function such that

\begin{itemize}
\item $K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$
\end{itemize}
We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.

@@

>>>>

Heymann et al. \cite{heymann} survey the field of spam on SNSs, identifying several common approaches. Identification-based approaches identify spam to train classifiers based on labels submitted by users or trusted moderators. Rank-based approaches demote visibility of questionable content, while interface-based approaches apply policies to prevent unwanted behavior. This work groups classification-based approaches with detection, although classifiers can be used in conjunction with user information to prevent spam before it happens.

@@

De Wang et al. \cite{wang} propose a cross-site spam detection framework to share spam data across all social networking sites, building classifiers to identify spam in profiles, messages and web pages. This multi-pronged approach lends itself to associative classification, in which, for example, a message would be classified as spam if it contained a link to a web page that had a high probability of being spam. Unfortunately, the differing characteristics of various social networks \eg the length of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora across diverse sites.

Facebook \cite{stein} provides an overview of their ``immune system" defences against phishing, fraud and spam. The system is composed of classifier services, an ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare features for classification and a policy engine to take action on suspected misuse. While the discussion remains high-level and includes few implemention particulars, it does include significant detail on the various types and characteristics of undesirable activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.

@@

Bosma et al. \cite{bosma} explore user-generated spam reports as a tool for building an unsupervised spam detection framework for SNSs. Their approach counts the number of spam reports against a suspected spammer and adds weight to reports based on user reputation. Determining reputation and trustworthiness of users in social networks has been well studied \cite{bian, guha, zhang} and appears to be a promising addition to social spam classification. The framework uses a Bayesian classifier and links messages with similar content, but does not take into account other features. Nevertheless, this is one of the few studies to test its framework on non-public data, including private messages, spam reports and user profiles from a large Dutch social networking site.

@@

For this project, we had unlimited access to data from InterPals, a SNS for users who wish to communicate with others from other countries, whether for language practice, cultural exchange or friendship. Users sign up by completing a registration form with information about themselves, including age, sex and location. After registering, users can expand their profile page to include descriptions, interests, languages they speak, etc. and upload photographs. After clicking on a link sent to their email address, a user can begin to interact with others on the site via private message, public ``wall" posts, comments on photos and on a bulletin board system.

@@

\textbf{Message body \& subject:} We extracted both the message body and subject from 
each spam and ham message. The cleaning script generated two additional features based 
on this. First, we counted the frequency of non-standard punctuation (we noticed that many 
spam messages would, for example, put spaces before commas, periods and quotation marks, 
and omit spaces after these characters). Second, we calculated the ratio of uppercase to 
lowercase letters in the text, after observing relatively high amounts of uppercase text in spam messages.
%% add numbers

\textbf{Recipient age, sex and country:} These fields were the age, sex and ISO 3166-1 alpha-2 code of the message 
recipient, as listed on their profile. Unfortunately, due to data collection issues, this data was only available 
for the most recent 11,000 spam messages. While the male-to-female ratio of the recipients was very close to that 
of non-spam recipients (both were 47\% male to 53\% female), we found that the age of recipients was typically higher.

\textbf{Recipient replied:} This boolean value indicates whether the recipient replied to the message. We saw 
that the mean reply rate for ham messages was 81.77\% (SD: 38.61), while it was only 0.81\% (SD: 8.97) for spam messages.

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
    \caption{Number of unique IP addresses associated with user account.}
    \label{fig:uniqip}
\end{figure}

@@

Linear regression is a method to model the relationship between an output variable, 
$y$, and explanatory variables, $X$. The output variable is a linear sum of the explanatory 
variables multiplied by their corresponding coeffecients, represented by $\beta$. 
In our case, $y$ represents a message's label as spam or ham (1 or 0), 
while $X$ is the feature vector, a bag-of-words representation of the message text and expert features. 
This implies that $\beta$ represents the weight of a feature's correlation to whether 
the message is spam or not. This gives us the simple regression formula: $$y=X\beta$$

In this project, we performed ordinary least squares regression. By minimizing the sum of squared 
residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution. 
$$\beta=(X'X)^{-1}X'y$$ In order to ensure that the matrix $X'X$ is invertible, we perform ridge regression. 
In ridge regression, we add the identity matrix, $I$, scaled by a factor, $\lambda$, to 
produce the following equation. $$\beta=(X'X+\lambda I)^{-1}X'y$$

We calculate this $\beta$ for a training set of bag-of-words features along with their corresponding labels. 
We then evaluate our model $\beta$ on the remaining messages in order to perform 10-fold cross-validation. 
We repeat this process with the set that includes expert features.

All of our linear regression code was implemented in MATLAB because of its ease-of-use 
and optimizations on matrix multiplications. We read in the text-based feature dictionaries and feature matrix
representations created above,

%We parse the messages from the text format produced by our cleaning script, using the dictionary generated earlier map words to integers, saving the feature data into two sparse matrices. The first matrix only includes the bag-of-words representation of the message body, while the second includes the expert features as well.  Matlab uses Compressed Sparse Column format, so we actually stored the matrix $X'$ because the dictionary is very sparse.

%For each feature set, the actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter. The multiplication $X'X$ takes up too much memory if we use the entire dictionary, so we chose to limit the dictionary size. We found that a dictionary of the 10,000 most frequent words gave sufficient features while also being computable using MATLAB's built-in sparse matrix multiplications. We also tried using a dictionary of 15,000 words, but only had a marginal improvement along with much slower execution. Using our 10,000 word dictionary, we calculated the performance of our matrix multiply to be 0.123 Gflops/s by dividing the number of non-zero elements in our matrices by the time to compute the multiplication. 
%
%
%Since we used ridge regression, we had the ability to tune the $\lambda$ parameter in order to get the best performance. As shown in figure \ref{fig:lambda}, we observed an AUC improvement of about .5\% between the default value of 1 and a tuned parameter of 1500. This is a fairly major difference since the AUC is at 93\% using the default value. Additionally, we also noticed a major difference in the top words when using better $\lambda$ values. Using the default value of 1, our top words included obscure words including ``\_c'', ``\_j'', ``donovan'', and ``matlock.'' When we used the tuned $\lambda$ value of 1500, the words were much more representative, as shown in table \ref{tab:words}.
%

@@

Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is a real-valued function such that: $$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.

@@

We tried a number of values for $\lambda$: 1, 10, 50, 100, 500, 1000, 1500, 3000. Figure \ref{fig:lambda}

\begin{tabular}{c|c
